Sleep Quality Estimation using Machine Learning and Thigh-Mounted Accelerometer Data in Free Living Conditions

### 2023-02-06 presentation/index.qmd
create hyperparameter tuning plot, ROC curve plots, and correct combined performance metrics

---
title: "Machine Learned Models for Estimation of Sleep Quality in Free-Living Accelerometer Data"
subtitle: "(working title)"
author: "Esben Lykke, PhD student"
date: "`r format(Sys.Date(), '%d %B, %Y')`"
format:
  revealjs:
    smaller: true
    theme: moon
    slide-number: true
    preview-links: auto
    logo: images/SDU_WHITE_RGB-png.png
    footer: https://github.com/esbenlykke/sleep_study
    css: my_styles.css
embed-resources: true
editor_options: 
  chunk_output_type: console
smooth-scroll: true
bibliography: references.bib
---

@@

```{r}
#| echo: false
library(gt)
library(arrow)
library(visNetwork)
library(tidymodels)
```

Lists can optionally be displayed incrementally:

::: incremental
-   First item
-   Second item
-   Third item
:::

. . .

Insert pauses to make other types of content display incrementally.

::: footer
Learn more: [Incremental Lists](https://quarto.org/docs/presentations/revealjs/#incremental-lists)
:::

@@

Overview
Network plot of data preparation steps

. . .

network plot

@@

::: {.column width="30%"}

@@

::: {.column width="40%"}

@@

## {.smaller transition="convex" .column}

```{r}
source("~/sleep_study/presentation/code/tables_performance.R")

@@

## {.smaller transition="convex" .column}

@@

```{r}
in_bed_tree <-
  read_rds("~/sleep_study/data/models/fitted_models/in_bed_simple_CART_fit.rds")

in_bed_tree |>
  extract_fit_engine() |>
  visTree(
    height = "550px",
    nodesPopSize = TRUE,
    minNodeSize = 10,
    maxNodeSize = 30,
    fallenLeaves = TRUE,
    legend = FALSE,
    digits = 3,
    direction = "UD",
    colorVar = c(
      "#CC5500", "#8B0000", "#FFD700",
      "#228B22", "#654321", "#8B0000", "#FFDB58"
    ),
    export = FALSE
  ) |>
  visNodes(font = list(
    color = "#EEE8D5",
    size = 20,
    strokeWidth = 1
  )) |>
  visEdges(font = list(
    color = "#EEE8D5",
    size = 20,
    strokeWidth = 1
  )) |> 
  visInteraction(dragNodes = FALSE, 
                 dragView = FALSE, 
                 zoomView = FALSE)
```

@@

```{r}
sleep_tree <-
  read_rds("~/sleep_study/data/models/fitted_models/sleep_simple_CART_fit.rds")

sleep_tree |>
  extract_fit_engine() |>
  visTree(
    height = "550px",
    nodesPopSize = TRUE,
    minNodeSize = 10,
    maxNodeSize = 30,
    fallenLeaves = TRUE,
    legend = FALSE,
    digits = 3,
    direction = "UD",
    colorVar = c(
      "#CC5500", "#8B0000", "#FFD700",
      "#228B22", "#654321", "#8B0000", "#FFDB58"
    ),
    export = FALSE
  ) |>
  visNodes(font = list(
    color = "#EEE8D5",
    strokeWidth = 1
  )) |>
  visEdges(font = list(
    color = "#EEE8D5",
    strokeWidth = 1
  )) |> 
  visInteraction(dragNodes = FALSE, 
                 dragView = FALSE, 
                 zoomView = FALSE)
```

>>>>

---
title: "Sleep Quality Estimation using Machine Learning and Thigh-Mounted Accelerometer Data in Free Living Conditions"
subtitle: "(working title)"
author: "Esben Lykke, PhD student"
date: "`r format(Sys.Date(), '%d %B, %Y')`"
format:
  revealjs:
    smaller: true
    theme: moon
    slide-number: true
    preview-links: auto
    logo: images/SDU_WHITE_RGB-png.png
    footer: https://github.com/esbenlykke/sleep_study
    css: my_styles.css
embed-resources: true
editor_options: 
  chunk_output_type: console
smooth-scroll: true
bibliography: references.bib
---

@@

```{r}
#| echo: false
library(gt)
library(arrow)
library(visNetwork)
library(tidymodels)
library(tidymodels)
```

- Sleep plays a vital role in health, thus, improving the assessment of sleep–wake outside of a laboratory environment is critical
- The gold standard (PSG) is costly and inconvenient.
- Methods for estimating sleep/wake based on accelerometry exist, primarily from wrist-worn devices
- Cole-Kripke and Sadeh algorithms are commonly used
- determine in-bed time is difficult, usually set by sleep log and/or human scorers
- detect wakefulness is difficult, worse performance in populations with sleep disorders
- typically a two level analysis: epoch based and summarized across night(s)
- Zmachine-derived sleep stats
-

Purpose...

## But Esben, what about them sleep stages!?

* I did free-living PSG recordings of sleep but...
  - Super fragile -> shitty data
  - Combersome and time consuming
  - free-living?
  - would surface skin temperature + acc be enough? Most likely needs HR

It was likely a dead end from the get-go :(

@@

data preparation, big time-consumer is handling raw acc data

only thigh data used. HSBC and other is only thigh data...

could be interesting to build models on thigh and hip ocmbined.

all zm recording is considered as in-bed (sensor problem?)

no sleep stages, only sleep/awake

sensor problems during sleep, up to 20 consecutive epochs (200 sec) are treated as sleep

@@

::: {.fragment .column width="30%"}

@@

::: {.fragment .column width="40%"}

@@

## building Models {.smaller transition="convex"}

```{r}
source(here::here("presentation/code/plot_hyperparameter_tuning.R"))

hyper_plot
```

## Results {transition="fade"}

::: {.fragment .column width="50%"}

* Performance Metrics
  - F1 Score
  - Accuracy
  - Sensitivity
  - Specificity
  - ROC curves

::: {.fragment .column width="50%"}

* Agreement With Zmachine Sleep Stats
  - Sleep Period Time
  - Total Sleep Time
  - Sleep Efficiency
  - Latency Until Persistent Sleep
  - Wake After Sleep Onset

## ROC Curves {.smaller transition="fade" .column-left}

```{r}
source(here::here("presentation/code/plot_roc_curves.R"))

roc_plots
```

## {.smaller transition="fade" .column-left}

```{r}
source(here::here("presentation/code/tables_performance.R"))

@@

## {.smaller transition="fade" .column}

@@

```{r}
in_bed_tree <-
  read_rds(here::here("data/models/fitted_models/in_bed_simple_CART_fit.rds"))

in_bed_tree |>
  extract_fit_engine() |>
  visTree(
    height = "550px",
    nodesPopSize = TRUE,
    minNodeSize = 10,
    maxNodeSize = 30,
    fallenLeaves = TRUE,
    legend = FALSE,
    digits = 3,
    direction = "UD",
    colorVar = c(
      "#CC5500", "#8B0000", "#FFD700",
      "#228B22", "#654321", "#8B0000", "#FFDB58"
    ),
    export = FALSE
  ) |>
  visNodes(font = list(
    color = "#EEE8D5",
    size = 20,
    strokeWidth = 1
  )) |>
  visEdges(font = list(
    color = "#EEE8D5",
    size = 20,
    strokeWidth = 1
  )) |>
  visInteraction(
    dragNodes = FALSE,
    dragView = FALSE,
    zoomView = FALSE
  )
```

@@

```{r}
sleep_tree <-
  read_rds(here::here("data/models/fitted_models/sleep_simple_CART_fit.rds"))

sleep_tree |>
  extract_fit_engine() |>
  visTree(
    height = "550px",
    nodesPopSize = TRUE,
    minNodeSize = 10,
    maxNodeSize = 30,
    fallenLeaves = TRUE,
    legend = FALSE,
    digits = 3,
    direction = "UD",
    colorVar = c(
      "#CC5500", "#8B0000", "#FFD700",
      "#228B22", "#654321", "#8B0000", "#FFDB58"
    ),
    export = FALSE
  ) |>
  visNodes(font = list(
    color = "#EEE8D5",
    strokeWidth = 1
  )) |>
  visEdges(font = list(
    color = "#EEE8D5",
    strokeWidth = 1
  )) |>
  visInteraction(
    dragNodes = FALSE,
    dragView = FALSE,
    zoomView = FALSE
  )
```

## Discussion {.smaller transition="zoom"}

- heteroscedasticity
- Cheung 2018 table 4: actigraphy provides a sufficiently narrow range of possible mean differences (CI 95%) clinical significant thresholds

## References {transition="convex"}
